\pkg{dataMaid}: your assistant for data cleaning in \proglang{R}

### 2017-01-03 latex/article_vol2.tex
Minor revisions of section 5.1

\subsection{Cleaning large datasets}
If the dataset becomes very large, the standard use of \R{clean()}
outlined above might not be ideal. If there is a vast number of
variables, creation of the \pkg{rmarkdown} document might be quite
slow, while a large number of observations will generally affect the
rendering time of the document. In this section, we give a few
practical examples of ways to deal with large data, while wishing to
still produce (potentially very long) data cleaning overview
documents. Note that the interactive tools of \pkg{dataMaid} can be used
as usual or sequentially in small subsets of the large dataset, if no
such overview documents are needed.

@@

\subsubsection{Economic memory use} 
Available memory may pose a problem with larger datasets ---
especially on machiens running Windows. The problem is, that older,
32-bit versions of Windows cannot necessary access all physical RAM,
while newer 64-bit versions of Windows cannot access virtual memory,
which makes \proglang{R} unnecessary slow.  Another solution, which is
especially relevant to Windows users due to %the unfortunate
the memory control for this operating system,
%combination of memory control in this operating system and
%\proglang{R},
%\hl{And also just R, right? Du har udkommenterert denne kommentar,
%men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
%det uden en præcision af hvad det går ud på. Hvad tænker du?}
is simply splitting the two steps performed
by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
afterwards. If the \R{rmarkdown} file is very long, as it will
typically be for very large datasets, keeping this file open in memory
wastes precious memory capacities. Therefore, we advice users to
instead split the two steps. This can be done in the following manner:

\begin{Schunk}
\begin{Sinput}
> clean(toyData, render = FALSE, openResult = FALSE}
> render("dataMaid_toyData.Rmd", quiet = FALSE)
\end{Sinput}
\end{Schunk}
This also deals with the fact that \pkg{dataMaid} can produce
\pkg{rmarkdown} files that supersedes the upper size limit of, say,
RStudio (which currently has an editor file size limit of 5MB). %\hl{dette giver
%ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 
%kommentar.}
% of RStudio,
%which is currently \hl{find number} GBs (using RStudio version
%1.0.44). 
%\hl{Is this maybe too editor specific? On the other hand, a
%  lot of people do use RStudio...}.

>>>>

\subsection{Cleaning large datasets}
If the dataset becomes very large, the standard use of \R{clean()}
outlined above might not be ideal. If there is a vast number of
variables, creation of the \pkg{rmarkdown} document might be quite
slow, while a large number of observations will generally affect the
rendering time of the document. This issue is especially relevant to 
Windows users, where memory usage is not very efficient. The problem is that older,
32-bit versions of Windows generally cannot access all physical RAM,
while newer 64-bit versions of Windows cannot access virtual memory. Therefore,
memory allocation issues becomes even more pressing when using this operating system.

In this section, we give a few
practical examples of ways to deal with large data, while wishing to
still produce (potentially very long) data cleaning overview
documents. Note that the interactive tools of \pkg{dataMaid} can be used
as usual or sequentially in small subsets of the large dataset, if no
such overview documents are needed.

@@

\subsubsection{Economic memory use} 
Another solution, which is
especially relevant to Windows users due to %the unfortunate
the memory handling on this operating system,
%combination of memory control in this operating system and
%\proglang{R},
is simply splitting the two steps performed
by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
afterwards. If the \R{rmarkdown} file is very long, as it will
typically be for very large datasets, keeping this file open in memory
wastes precious memory capacities. Therefore, we advice users to
instead split the two steps. This can be done in the following manner:

\begin{Schunk}
\begin{Sinput}
> clean(toyData, render = FALSE, openResult = FALSE}
> render("dataMaid_toyData.Rmd", quiet = FALSE)
\end{Sinput}
\end{Schunk}
This also deals with the fact that \pkg{dataMaid} can produce
\pkg{rmarkdown} files that supersedes the upper size limit of code editors, e.g.
RStudio, which currently has an editor file size limit of 5MB. %\hl{dette giver
%ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 
%kommentar.}
% of RStudio,
%which is currently \hl{find number} GBs (using RStudio version
%1.0.44). 
%\hl{Is this maybe too editor specific? On the other hand, a
%  lot of people do use RStudio...}.
